---
title: |
  <div class="tight-title">
    <div>ECOM30004/90004</div>
    <div>Time Series Analysis and Forecasting</div>
     <div class="spacer"></div> <!-- blank line -->
    <div class="bold">Tutorial 6</div>

    <div class="spacer"></div> <!-- blank line -->

    <div class="author">Tutor: Pengyun (Kevin) Chen</div>
    <div class="author">University of Melbourne</div>
    <div class="author">Semester 2, 2026</div>
  </div>

  
format:
  revealjs:
    theme: simple
    self-contained: true
    chalkboard: false
    width: 1300
    height: 950
    footer: 'Access Slides: `pengyunc.github.io/TSAF_Week6`'
    title-slide-attributes:
      data-background-color: "#000F46"
    slide-number: true
    center-title-slide: true
    toc: false
    css: styles.css
    callout-appearance: simple
    background-transition: fade
    callout-icon: true
    transition: slide
    scrollable: true
    code-block-height: 800px
    code-line-numbers: true
    mathjax: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-svg.js"
    html-math-method: mathjax
    execute:
      echo: true
      eval: true
      enabled: true
---

## Notice

::: {.callout-warning title="Conditions of Use"}

- This document is intended for personal use and is NOT meant to be shared on the internet (e.g. Studocu, ThinkSwap, StudentVIP, Coursehero, etc.)
- Do not share this document with anyone, even if they are enrolled in ECOM30004/90004.
- Official course materials, such as lecture slides, tutorial sheets and prescribed readings should always be your primary point of reference.

:::
<div style="height: 0.2em;"></div>
Slides accessible here: `pengyunc.github.io/TSAF_Week6`


## This tutorial
<blue>In this tutorial, we demonstrate forecasting the 2020Q1 CPI level, including stationarity tests, ARIMA model selection, and the final forecast computation.</blue>

<div style="height: 0.2em;"></div>
:::{.callout-note title = "Agenda"}

- <blue>Exercise 1(a):</blue> Transform CPI levels into log differences and select the ADF specification (lags).

- <blue>Exercise 1(b):</blue> Test the log-differenced series for a unit root (ADF).

- <blue>Exercise 1(c):</blue> Select the ARIMA model using AICc and Ljung–Box diagnostics.

- <blue>Exercise 1(d):</blue> Forecast 2020Q1 inflation and the CPI level.

- <blue>Exercise 2:</blue> Explain why first differencing removes a linear trend.
:::

<div style="height: 0.2em;"></div>



## We know that taking the logs is not enough... {auto-animate="true"}
Recall the unit root test results for log `CPI` from Self-directed Activities 5.
```{r}
dt <- read.csv("CPI.csv")
Y <- ts(log(dt$CPI), start = c(1922,2), end = c(2025,2), frequency=4)
Y <- window(Y, start = c(1946,1), end = c(2019,4))
```

<div class="horizontal-line"></div>
```{r}
plot(Y)
```

An upward trend is obvious <blue>$\to$ add a time trend in deterministic specification<blue>

## We know that taking the logs is not enough... {auto-animate="true"}
:::{.callout-question}
Carry out an ADF test of `Y`, present (i) the estimated test equation, 
(ii) the $t_{\text{ADF}}$ test statistic, 5\% critical value and decision, and (iii) its $p$-value.
:::
<div style="height: 0.5em;"></div>

```{r}
library(urca)    # For use of ur.df function
adf <- ur.df(Y, type = "trend", selectlags = "AIC")
p_adf <- punitroot(adf@teststat[1], trend = "ct", statistic = "t")
```
<div style="height: 0.5em;"></div>
::: {.panel-tabset}

## (i) test equation
The test equation is:
$$\Delta Y_{t}=\beta_{0}+\beta_{1}t+\varphi Y_{t-1}+\sum^{p-1}_{k=1}\psi_{k} \Delta Y_{t-k}+U_{t}$$
The hypotheses are:
\begin{align}
H_{0}:&\;\varphi=0& &\text{($Y_{t}$ has a unit root)}\\
H_{1}:&\;\varphi<0& &\text{($Y_{t}$ does not have a unit root)}
\end{align}
The number of lags of $\Delta Y_{t}$ (i.e., $p-1$) is selected by an information criterion (AIC), where $p$ is the order of the underlying AR model.

## (ii) test summary
```{r}
summary(adf)
```

## (iii) $p$-value
```{r}
print(p_adf)
```
<div style="height: 0.5em;"></div>
Since $p=0.997>0.05$, we <teal>**fail to reject** the null hypothesis $H_{0}$: there is unit root</teal>.
<div style="height: 0.5em;"></div>

<blue>$\implies$ Simply removing the trend component is not enough to make $Y_{t}$ stationary. </blue>

<blue>$\implies$ We need to take the first difference of log `CPI`.</blue>
:::

## Exercise 1(a)
:::{.callout-question}
Next create the first difference time series $\Delta Y_{t}$ and a time series plot. Decide whether a trend is required for a unit root test for $\Delta Y_{t}$. 
:::

<div style="height: 0.5em;"></div>

```{r}
DY <- diff(Y)   # DY is the log difference of the CPI -> Inflation
plot(DY)
```
<blue>**Observation:** There is no evidence of a linear trend in this time series.</blue> There are some extended periods (e.g. early 1950's and 1970-80's) when inflation was higher.

## Exercise 1(b)  {auto-animate="true"}
:::{.callout-question}
Carry out the unit root test on $\Delta Y_{t}$. Conclude how many differences are required for log CPI to be stationary.
:::

<div style="height: 0.5em;"></div>

```{r}
adf <- ur.df(DY, type = "drift", selectlags = "AIC")
p_adf <- punitroot(adf@teststat[1], trend = "c", statistic = "t")
summary(adf)
```
<div class="horizontal-line"></div>
The estimated ADF test equation is:

$$\widehat{\Delta(\Delta Y_{t})}=0.0031-0.258\Delta Y_{t-1}-0.4253\Delta(\Delta Y_{t-1})$$

## Exercise 1(b)  {auto-animate="true"}
:::{.callout-question}
Carry out the unit root test on $\Delta Y_{t}$. Conclude how many differences are required for log CPI to be stationary.
:::

<div style="height: 0.5em;"></div>

```{r}
adf <- ur.df(DY, type = "drift", selectlags = "AIC")
p_adf <- punitroot(adf@teststat[1], trend = "c", statistic = "t")
summary(adf)
```
<div class="horizontal-line"></div>
The $t$-statistic is $-5.1529$, and the $5\%$ critical value is $-2.87$.

<blue>Since $-5.1539<-2.87$, we **reject** $H_0$: $\Delta Y_{t}$ has a unit root.</blue>


## Exercise 1(b)  {auto-animate="true"}
:::{.callout-question}
Carry out the unit root test on $\Delta Y_{t}$. Conclude how many differences are required for log CPI to be stationary.
:::

<div style="height: 0.5em;"></div>

```{r eval=FALSE}
adf <- ur.df(DY, type = "drift", selectlags = "AIC")
p_adf <- punitroot(adf@teststat[1], trend = "c", statistic = "t")
summary(adf)
```
<div style="height: 0.5em;"></div>

The $t$-statistic is $-5.1529$, and the $5\%$ critical value is $-2.87$.

Since $-5.1539<-2.87$, we **reject** $H_0$: $\Delta Y_{t}$ has a unit root.

::: {.fragment}
<div class="horizontal-line"></div>
```{r}
print(p_adf)
```
<div style="height: 0.5em;"></div>
<teal>Alternatively, since $p=9.82\times 10^{-6}<0.05$, we **reject $H_0$: $\Delta Y_{t}$ has a unit root**.</teal>
:::

:::{.fragment}
<div class="horizontal-line"></div>
<blue>**Conclusion:** To render log CPI stationary, it is necessary to difference the series *once*.</blue>
:::


## Exercise 1(c)  {auto-animate="true"}
:::{.callout-question}
Specify and estimate an appropriate ARIMA($p,d,q$) model for $Y_{t}$.
:::
<div style="height: 0.5em;"></div>
:::{.callout-tip}
- Since first differencing is required for log CPI to be stationary, we need $d=1$.
- How to select $(p,q)$? <green>$\to$ Conduct a grid search for $p=0,\dots,4$ and $q=0,\dots,4$</green>
:::

<div class="horizontal-line"></div>
:::{.callout-note title="Code Structure"}
- Set up container matrices for AICc values (`AICc`) and Ljung-Box test $p$-values (`LBp`)

- For each $(p,q)$ combination:
  - Fit the ARIMA($p,1,q$) model with `Arima()` function, and extract AICc
  
  - Extract the Ljung-Box $p$-values with `checkresiduals()` function
  
- Choose the model with the **lowest** AICc, and verify that its <teal>Ljung-Box $H_{0}$ (no autocorrelation)</teal> is not rejected.

:::


## Exercise 1(c)  {auto-animate="true"}
:::{.callout-question}
Specify and estimate an appropriate ARIMA($p,d,q$) model for $Y_{t}$.
:::
<div style="height: 0.5em;"></div>
:::{.callout-tip}
- Since first differencing is required for log CPI to be stationary, we need $d=1$.
- How to select $(p,q)$? <green>$\to$ Conduct a grid search for $p=0,\dots,4$ and $q=0,\dots,4$</green>
:::


```{r}
library(forecast)

pmax <- 4
qmax <- 4
LBp <- matrix(nrow=pmax+1, ncol=qmax+1)
rownames(LBp) <- paste0("p=", 0:pmax)
colnames(LBp) <- paste0("q=", 0:qmax)
AICc <- LBp

for (p in 0:pmax){
  for (q in 0:qmax){
    eq <- Arima(DY, order=c(p,0,q))
    capture.output(LBp[p+1,q+1] <- checkresiduals(eq,
      plot=FALSE)$p.value)
    AICc[p+1,q+1] <- eq$aicc
  }  
}
```


## Exercise 1(c)  {auto-animate="true"}
:::{.callout-question}
Specify and estimate an appropriate ARIMA($p,d,q$) model for $Y_{t}$.
:::
<div style="height: 0.5em;"></div>

Residual autocorrelation result (`=0` if $H_{0}$ maintained)
```{r}
print(1*(LBp<0.05))
```
<div style="height: 0.5em;"></div>

Model to minimise AICc:
```{r}
print(AICc==min(AICc))
```
<div style="height: 0.5em;"></div>

:::{.fragment}

<blue>**Conclusion:** The chosen model is ARIMA($1,1,3$) for $Y_{t}$.</blue>
:::


## Exercise 1(d)  {auto-animate="true"}
:::{.callout-question}
<blue>Use the chosen model to calculate a forecast of inflation</blue> and the CPI for the March quarter of 2020.
:::
<div style="height: 0.5em;"></div>
:::{.callout-tip}
Note that we have selected ARMA(1,3) for $\Delta Y_{t}\equiv \Delta\log(CPI_{t})$ in Exercise 1(c).
:::

:::{.fragment}
<div style="height: 0.1em;"></div>
Forecast 2020Q1 inflation:  $\Delta{}Y_{2020Q1}=\log(CPI_{2020Q1})-\log(CPI_{2019Q4})$
```{r}
eq <- Arima(DY, order = c(1,0,3))
ForecastInflation_2020q1 <- forecast(eq, h=1)$mean
print(ForecastInflation_2020q1)
```
:::

:::{.fragment}
<div class="horizontal-line"></div>
**The inflation forecast is based on an approximation…**

Recall from Week 1 tutorial that for small $r$ values, $\log(1+r)\approx r$:
\begin{align}
CPI_{2019Q4}(1+\text{inflation}_{2020Q1})&=CPI_{2020Q1}\\
\log(1+\text{inflation}_{2020Q1})&=\log(CPI_{2020Q1})-\log(CPI_{2019Q4})\\
\implies\qquad \color{blue}{\text{inflation}_{2020Q1}}&\approx\log(CPI_{2020Q1})-\log(CPI_{2019Q4})
\end{align}
:::


## Exercise 1(d)  {auto-animate="true"}
:::{.callout-question}
Use the chosen model to calculate a forecast of inflation and <blue>the CPI for the March quarter of 2020</blue>.
:::
<div style="height: 0.5em;"></div>
:::{.callout-tip}
Note that we have selected ARMA(1,3) for $\Delta Y_{t}\equiv \Delta\log(CPI_{t})$ in Exercise 1(c).
:::
<div style="height: 0.1em;"></div>
**How can we forecast the 2020Q1 CPI?**

We know that
\begin{align}
\color{blue}{\text{inflation}_{2020Q1}}&\approx\log(CPI_{2020Q1})-\log(CPI_{2019Q4})
\end{align}

:::{.fragment}
thus we have $\widehat{CPI}_{2020Q1}=CPI_{2019Q4}\times(1+\widehat{\operatorname{inflation}}_{2020Q1})$

```{r}
# Create a ts of CPI (i.e. not log CPI)
CPI <- ts(dt$CPI, start=c(1922,2), end=c(2025,2), frequency=4)
CPI_2019q4 <- CPI[which(time(CPI)==2019.75)]

# Forecast CPI for 2020q1
ForecastCPI_2020q1 <- CPI_2019q4*(1+ForecastInflation_2020q1)
print(ForecastCPI_2020q1)
```
:::


## Exercise 1(d)  {auto-animate="true"}
:::{.callout-question}
Use the chosen model to calculate a forecast of inflation and <blue>the CPI for the March quarter of 2020</blue>.
:::
<div style="height: 0.5em;"></div>

**How can we forecast the 2020Q1 CPI?**

First approach: $\widehat{CPI}_{2020Q1}=CPI_{2019Q4}\times(1+\widehat{\operatorname{inflation}}_{2020Q1})$
```{r echo=FALSE}
print(ForecastCPI_2020q1)
```
<div class="horizontal-line"></div>

:::{.fragment}
<blue>Second approach:</blue>

- Since we know that <blue>$\widehat{\log(CPI_{2020Q1})}=\log(CPI_{2019Q4})+\widehat{\text{inflation}}_{2020Q1}$</blue>
- then we can just take the exponential of the logs:
<blue>$$\widehat{CPI_{2020Q1}}=\exp(\widehat{\log(CPI_{2020Q1})})=\exp(\log(CPI_{2019Q4})+\widehat{\text{inflation}}_{2020Q1})$$</blue>
:::


## Exercise 1(d)  {auto-animate="true"}
:::{.callout-question}
Use the chosen model to calculate a forecast of inflation and <blue>the CPI for the March quarter of 2020</blue>.
:::
<div style="height: 0.5em;"></div>
**How can we forecast the 2020Q1 CPI?**

First approach: $\widehat{CPI}_{2020Q1}=CPI_{2019Q4}\times(1+\widehat{\operatorname{inflation}}_{2020Q1})$
```{r echo=FALSE}
print(ForecastCPI_2020q1)
```
<div class="horizontal-line"></div>

<blue>Second approach: $\widehat{CPI_{2020Q1}}=\exp(\log(CPI_{2019Q4})+\widehat{\text{inflation}}_{2020Q1})$</blue>

```{r}
n <- length(Y)
ForecastY_2020Q1 <- Y[n] + ForecastInflation_2020q1
print(ForecastY_2020Q1)      # Forecasted log CPI at 2020Q1
```
<div style="height: 0.3em;"></div>
```{r}
ForecastCPI_2020q1 <- exp(ForecastY_2020Q1)
print(ForecastCPI_2020q1)    # Forecasted CPI level at 2020Q1
```


## Exercise 1(d)  {auto-animate="true"}
:::{.callout-question}
Use the chosen model to calculate a forecast of inflation and <blue>the CPI for the March quarter of 2020</blue>.
:::
<div style="height: 0.5em;"></div>

:::{.callout-tip}
Recall that we modeled $\Delta Y_{t}$ as ARMA(1,3) <green>$\Rightarrow$ $Y_{t}$ should be modeled as ARMA(1,1,3)</green>.
:::

::: {.panel-tabset}

## ARIMA(1,1,3) for $Y_{t}\equiv\log(CPI_{t})$

```{r}
eq1 = Arima(Y, order = c(1,1,3), xreg = (1:n))
print(eq1)
```
<div style="height: 0.1em;"></div>
$\displaystyle\textsf{The model is:}\quad\boxed{\quad\begin{aligned}
\color{blue}{Y_{t}}\,&\color{blue}{=\beta_{1}t+Z_{t}}\\
\Delta Z_{t}&=\phi_{1}\Delta Z_{t-1}+U_{t}+\theta_{1}U_{t-1}+\theta_{2}U_{t-2}+\theta_{3}U_{t-3}
\end{aligned}\quad}$

::: {.absolute top="530px" left="820px" style="font-size: 36px;"}
<blue>These are two versions of the same model and give equivalent coefficient estimates!</blue>
:::



## ARMA(1,3) for $\Delta Y_{t}\equiv\Delta \log(CPI_{t})$
```{r}
eq = Arima(DY, order = c(1,0,3))
print(eq)
```
$\displaystyle\textsf{The model is:}\quad\boxed{\quad\begin{aligned}
\color{blue}{\Delta Y_{t}}\,&\color{blue}{=\beta_{1}+\Delta Z_{t}}\\
\Delta Z_{t}&=\phi_{1}\Delta Z_{t-1}+U_{t}+\theta_{1}U_{t-1}+\theta_{2}U_{t-2}+\theta_{3}U_{t-3}
\end{aligned}\quad}$
:::

::: {.absolute top="530px" left="820px" style="font-size: 36px;"}
<blue>These are two versions of the same model and give equivalent coefficient estimates!</blue>
:::



## Exercise 1(d)  {auto-animate="true"}
:::{.callout-question}
Use the chosen model to calculate a forecast of inflation and <blue>the CPI for the March quarter of 2020</blue>.
:::
<div style="height: 0.5em;"></div>

:::{.callout-tip}
Recall that we modeled $\Delta Y_{t}$ as ARMA(1,3) <green>$\Rightarrow$ $Y_{t}$ should be modeled as ARMA(1,1,3)</green>.
:::
<div style="height: 0.1em;"></div>
We can forecast $Y_{2020Q1}\equiv\log(CPI_{2020Q1})$ with <purple>ARMA(1,3) of $\Delta Y_{t}$ with drift</purple>:

```{r}
eq = Arima(DY, order = c(1,0,3))
ForecastInflation_2020q1 <- forecast(eq, h=1)$mean
n <- length(Y)
ForecastY_2020Q1 <- Y[n] + ForecastInflation_2020q1
print(ForecastY_2020Q1)
```

:::{.fragment}
<div style="height: 0.8em;"></div>
We can also forecast $Y_{2020Q1}$ with <teal>ARIMA(1,1,3) of $Y_{t}$ with drift and trend</teal>:
```{r}
eq1 = Arima(Y, order = c(1,1,3), xreg = (1:n))
ForecastY_2020q1 <- forecast(eq1, xreg=(n+1))$mean
print(ForecastY_2020q1)
```
:::

:::{.fragment}

::: {.absolute top="548px" left="250px" style="font-size: 36px;"}
<blue>$\to\;\;\boxed{\;\textsf{Using either model is fine.}\;}$</blue>
:::

::: {.absolute top="833px" left="250px" style="font-size: 36px;"}
<blue>$\to\;\;\boxed{\;\textsf{Using either model is fine.}\;}$</blue>
:::

:::


## Exercise 1(d)  {auto-animate="true"}
:::{.callout-question}
Use the chosen model to calculate a forecast of inflation and <blue>the CPI for the March quarter of 2020</blue>.
:::
<div style="height: 0.5em;"></div>
**Remarks on the CPI forecasts:** 

Both approaches are approximations.

:::{.fragment}
<div class="horizontal-line"></div>
First approach: $\widehat{CPI}_{2020Q1}=CPI_{2019Q4}\times (1+\color{purple}{\underbrace{\widehat{\text{inflation}}_{2020Q1}}_{\text{approximation}}})$
:::

:::{.fragment}
<div class="horizontal-line"></div>
Second approach: $\widehat{CPI}_{2020Q1}=\exp(\log(\widehat{CPI}_{2020Q1}))$
:::




## Exercise 1(d)  {auto-animate="true"}
:::{.callout-question}
Use the chosen model to calculate a forecast of inflation and <blue>the CPI for the March quarter of 2020</blue>.
:::
<div style="height: 0.5em;"></div>
**Remarks on the CPI forecasts:** 

Both approaches are approximations.


<div class="horizontal-line"></div>
First approach: $\widehat{CPI}_{2020Q1}=CPI_{2019Q4}\times (1+\color{purple}{\underbrace{\widehat{\text{inflation}}_{2020Q1}}_{\text{approximation}}})$


<div class="horizontal-line"></div>
Second approach: $\widehat{CPI}_{2020Q1}=\exp(\color{teal}{\log(\widehat{CPI}_{2020Q1})}\color{black}{)}$

<teal>
<div style="height: 0.1em;"></div>
The inner part is an estimate of an expectation conditional on information set $\mathcal{Y}_{2019Q4}$:
$$\widehat{\log(CPI_{2020Q1})}\equiv\hat{Y}_{2020Q1}=\hat{\mathbb{E}}[Y_{2020Q1}\mid \mathcal{Y}_{2019Q4}]$$
</teal>

:::{.fragment}
$$\implies \widehat{\log(CPI_{2020Q1})}=\hat{\mathbb{E}}[\log(CPI_{2020Q1})\mid \mathcal{Y}_{2019Q4}]$$
and this is estimated through the ARIMA(1,1,3) model fitted on $Y_{t}$.
:::



## Exercise 1(d)  {auto-animate="true"}
:::{.callout-question}
Use the chosen model to calculate a forecast of inflation and <blue>the CPI for the March quarter of 2020</blue>.
:::
<div style="height: 0.5em;"></div>
**Remarks on the CPI forecasts:** 

Both approaches are approximations.


<div class="horizontal-line"></div>
First approach: $\widehat{CPI}_{2020Q1}=CPI_{2019Q4}\times (1+\color{purple}{\underbrace{\widehat{\text{inflation}}_{2020Q1}}_{\text{approximation}}})$


<div class="horizontal-line"></div>
Second approach: $\widehat{CPI}_{2020Q1}=\exp(\color{blue}{\hat{\mathbb{E}}[\log(CPI_{2020Q1})\mid\mathcal{Y}_{2019Q4}]}\color{black}{)}$
<div style="height: 0.5em;"></div>
<oblue>
**Question:** Does the following hold?
$$\color{blue}{\mathbb{E}[{CPI}_{2020Q1}\mid\mathcal{Y}_{2019Q4}]\overset{?}{=}\exp(\mathbb{E}[\log(CPI_{2020Q1})\mid \mathcal{Y}_{2019Q4}])}$$
</oblue>


:::{.fragment}
<blue>
**Answer:** No. This is because of Jensen's inequality - since $\exp(x)$ is convex:
$$\mathbb{E}[\exp(\log(CPI_{2020Q1})\mid \mathcal{Y}_{2019Q4})]>\exp(\mathbb{E}[\log(CPI_{2020Q1})\mid \mathcal{Y}_{2019Q4}])$$
</blue>
:::


## Exercise 1(d)  {auto-animate="true"}
:::{.callout-question}
Use the chosen model to calculate a forecast of inflation and <blue>the CPI for the March quarter of 2020</blue>.
:::
<div style="height: 0.5em;"></div>
**Remarks on the CPI forecasts:** 

Both approaches are approximations.


<div class="horizontal-line"></div>
First approach: $\widehat{CPI}_{2020Q1}=CPI_{2019Q4}\times (1+\color{purple}{\underbrace{\widehat{\text{inflation}}_{2020Q1}}_{\text{approximation}}})$


<div class="horizontal-line"></div>
Second approach: $\widehat{CPI}_{2020Q1}=\exp(\hat{\mathbb{E}}[\log(CPI_{2020Q1})\mid\mathcal{Y}_{2019Q4}]\color{black}{)}$
<div style="height: 0.5em;"></div>
<blue>
Thus this estimator is actually biased for the conditional mean  $\mathbb{E}[CPI_{2020Q1}\mid\mathcal{Y}_{2019Q4}]$:
$$\mathbb{E}[CPI_{2020Q1}\mid\mathcal{Y}_{2019Q4}]>\exp(\mathbb{E}[\log(CPI_{2020Q1})\mid \mathcal{Y}_{2019Q4}])$$
</blue>

:::{.fragment}
<green>
Nevertheless, in practice, simply taking the exponential tends to work well in providing a good approximation.
</green>
:::

## Exercise 2  {auto-animate="true"}
:::{.callout-question}
It seems like if $Y_{t}$ has a linear trend then $\Delta Y_{t}$ does not. Explain why this is.
:::
<div style="height: 0.5em;"></div>
:::{.callout-note}
<blue>Recall from Exercise 1(d):</blue> ARIMA(1,1,3) with trend on $Y_{t}$ produced the same parameter estimates as ARMA(1,3) with drift on $\Delta Y_{t}$. Why is that?
:::
<div style="height: 0.5em;"></div>

:::{.fragment}
The `Arima` function is estimating the following set of equations:
<div style="height: 0.5em;"></div>
:::

::: {.incremental}
- ARIMA(1,1,3) with drift and trend for $Y_{t}$:
\begin{align*}
Y_{t}&= \beta_{0}+\beta_{1}t+Z_{t}\\
\Delta Z_{t}&=\phi_{1}\Delta Z_{t-1}+U_{t}+\theta_{t} U_{t-1}+\theta_{2}U_{t-2}+\theta_{3}U_{t-3}
\end{align*}<div style="height: 0.5em;"></div>

- ARMA(1,3) with drift for $\Delta Y_{t}$:
\begin{align*}
\Delta Y_{t}&= \alpha+W_{t}\\
W_{t}&=\phi_{1}W_{t-1}+U_{t}+\theta_{t} U_{t-1}+\theta_{2}U_{t-2}+\theta_{3}U_{t-3}
\end{align*}

:::

:::{.fragment}
If both gives the exact same estimates, then $\alpha=\beta_{1}$ and $W_{t}=\Delta Z_{t}$.
:::


## Exercise 2 {auto-animate="true"}
:::{.callout-question}
It seems like if $Y_{t}$ has a linear trend then $\Delta Y_{t}$ does not. Explain why this is.
:::
<div style="height: 0.5em;"></div>
:::{.callout-note}
<blue>Show that $\alpha=\beta_{1}$, $W_{t}=\Delta Z_{t}$ in the following equations:</blue>

$$Y_{t}=\beta_{0}+\beta_{1}t+Z_{t}\qquad\qquad\Delta Y_{t}=\alpha+W_{t}$$
:::
<div style="height: 0.5em;"></div>

:::{.fragment}
Subtract $Y_{t-1}$ on both sides gives:

$$Y_{t}\,\color{blue}{-\,Y_{t-1}}\color{black}{\,=(\beta_{0}+\beta_{1}t+Z_{t})}\color{blue}{\,-\,(\beta_{0}+\beta_{1}(t-1)+Z_{t-1})}$$
:::

:::{.fragment}
Then simplify:
\begin{align*}

Y_{t}-Y_{t-1}&=\beta_{0}-\beta_{0}+\beta_{1}(t-(t-1))+Z_{t}-Z_{t-1}\\
\Delta Y_{t}&=\beta_{1}+\Delta Z_{t}
\end{align*}
:::

:::{.fragment}
We have thereby shown that $\alpha=\beta_{1}$ and $W_{t}=\Delta Z_{t}$.
<div style="height: 0.5em;"></div>
:::

:::{.fragment}
<blue>In addition, we have shown that taking the first difference removes the linear trend.</blue>
<blue>
$$Y_{t}=\beta_{0}+\beta_{1}t+Z_{t}\quad\xrightarrow{\;\textsf{take first difference}\;}\quad \Delta Y_{t}=\beta_{1}+\Delta Z_{t}$$
</blue>
:::


## {data-background-color="#000F46"}
<div style="height: 85vh; display: flex; align-items: center; justify-content: center; color: white; font-size: 2em;">
Thank you for coming!
</div>


<script>
document.querySelectorAll("myimage").forEach(el => {
  const src = el.getAttribute("src");
  const width = el.getAttribute("width") || "100%";

  const img = document.createElement("img");
  img.src = src;
  img.style.width = width;
  img.style.maxWidth = "100%";  // prevent overflow
  el.style.display = "block";
  el.style.textAlign = "center";
  
  el.appendChild(img);
});
</script>